Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-06-02 doc/paper.tex
fixed non-ascii in references (grr, only incompatible once \zh is used) and revised introduction somewhat

\begin{abstract}
An increased focus on a

is proposed.

@@

In this paper, we propose bringing linguists back from a premature exile to craft high-quality, ``designer'' synthetic parallel corpora for machine translation.
Such a division of labor would let linguists be linguists, with MT system development proceeding in parallel.  
In a sense, the use of external corpora has ``outsourced'' the rule-based MT linguist's job to third-party translators, who implicitly encode linguistic information in parallel corpora.  
What we propose is to bring these jobs back ``in-house'', in the updated context of corpus-based MT.

@@

% hmm, this would up being disturbingly short...
\section{Related Work}
\label{sec:related}

@@

Our parallel corpus generation system (Section \ref{sec:generator}) is very similar to the recursive equivalence class approach to example-based MT \mycitep{ebmtbook} used by \mynewcite{brown:99}. 
Our generator also uses a somewhat more varied grammar that also admits the use of participial noun modifiers.
But more importantly, we believe this line of work was ahead of its time and consider our main contribution to be bringing attention to its potential value to the contemporary machine learning approach to MT.

@@

%%%%% WAIT a second, doesn't this in principle just leave it up to Moses to pick up ``class phrases''?

%% uh...

@@

% let's figure out how to do a ``figure''

As a simple example, 
{\small \tt The NP jumped over the moon}

@@

Unlike AI-oriented or natural language understanding approaches, our primary goal is to generate parallel corpora of surface forms, offering us a significant amount of flexibility on the method of generation. 
adopt elements of both rule-based and example-based MT systems

@@

We adopt the pragmatic, ``defeatist'' attitude of modeling only straightforward linguistic phenomena compositionally; if the theory and data structures required to model a phenomenon are sufficiently complicated, everything is probably better off just going directly to a translation via template, instead of trying to concoct a sophisticated theory that has to be consistent with all previously theorized phenomena.
The tradeoff is that 
Compositional trees are more automatic-like, but generate more sentence pairs/tuples.
Complex templates are more hand-crafted, but generate fewer pairs/tuples, and are (presumably) harder to combine compositionally (but are they really!?).

@@

@@

@@

%RNN encoder-decoders are not the first corpus-based MT methodology (EBMT then SMT then this), and it is unlikely in this era of Big Data that they will be the last.

@@

% these are so short that it's probably better to have it in literally instead of using \input
\if \buildtarget \buildamta    
    \small
    \bibliographystyle{apalike}
\else \if \buildtarget \buildemnlp
    %Note that emnlp2016.bst chokes if there are 0 citations in your paper.    
    %\mynewcite{moses}    
    \bibliographystyle{emnlp2016}

>>>>

\begin{abstract}
An increased focus on a
S \zh{我把} O V
is proposed.

@@

One of the major advantages of data-driven MT is that it removes the need for linguists to develop a new set of rules for an MT system to support a new language.
In a sense, though, the rule-based MT linguist's job has really been  ``outsourced'' to third-party translators, who implicitly encode linguistic information in parallel corpora.
We believe there is still a significant ``in-house'' role for linguists to play in corpus-based machine translation, and that is to direct the crafting of high-quality, ``designer'' synthetic parallel corpora, leveraging technologies like crowdsourcing to maximize throughput.
Such a division of labor would let linguists be linguists, with MT system development proceeding in parallel.

@@

% hmm, this wound up being disturbingly short...
\section{Related Work}
\label{sec:related}

@@

Our parallel corpus generation system (Section \ref{sec:generator}) is very similar to the recursive equivalence class approach to example-based MT \mycitep{ebmtbook} used by \mynewcite{brown:99}. 
Our generator has more of a rule-based flavor, using a somewhat more varied grammar that also admits the use of participial noun modifiers.
But more importantly, we believe this line of work was ahead of its time and consider our main contribution to be bringing attention to its potential value to the contemporary machine learning approach to MT.

@@

@@

Unlike AI-oriented or natural language understanding approaches, our primary goal is to generate parallel corpora of surface forms, offering us a significant amount of flexibility as to the method of generation.

\small 
\begin{center}
{\tt S V O .}

%S \zh{把} O V

{\tt S \zh{我把} O V}

%\zh{把} {\tt O V} \zh{}
%\\
%S 
\end{center}
\normalsize

We adopt a pragmatic attitude of only trying to model straightforward linguistic phenomena compositionally; if the theory and data structures required to model a linguistic phenomenon are sufficiently complicated,

everything is probably better off just going directly to a translation via template, instead of trying to concoct a sophisticated theory that has to be consistent with all previously theorized phenomena.

At the time of this writing, our system can 
Literal 
sentence templates are all

Our parallel corpus generator starts with

native speakers, linguists or otherwise

S and O can be expanded

We can then allow the generator to vary both the syntactic and lexical content of

@@

adopt elements of both rule-based and example-based MT systems

@@

The tradeoff is that 
Compositional trees are more automatic-like, but generate more sentence pairs/tuples.
Complex templates are more hand-crafted, but generate fewer pairs/tuples, and are (presumably) harder to combine compositionally (but are they really!?).

@@

we of 
the {\small \tt tmcombine} script \mycitep{tmcombine} included with Moses

@@

Sufficient quantities of supervised data seem to underlie high or state of the art performance in NLP

@@

RNN encoder-decoders are not the first corpus-based MT methodology (EBMT then SMT then this), and it is unlikely in this era of Big Data that they will be the last.

@@

% these are so short that it's probably better to have it in literally instead of using \input
\if \buildtarget \buildamta    
    \small
    \bibliographystyle{apalike}
\else \if \buildtarget \buildemnlp
    %Note that emnlp2016.bst chokes if there are 0 citations in your paper.    
    \mynewcite{moses}    
    \bibliographystyle{emnlp2016}
